---
title: "Assignment 3: Apple Financial Data Analysis"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---
### Name: Chen Yongquan (G2002341D)
```{r}
library(TSA)
library(astsa)
library(fBasics)
library(forecast)
library(ggplot2)
library(quantmod)
library(rugarch)
library(tseries)
library(zoo)

AAPL = getSymbols("AAPL", from='2002-02-01', to='2017-02-01', src='yahoo', auto.assign = F) 
write.zoo(AAPL, file = "AAPL.csv", sep = ",", quote = FALSE)
```
```{r}
# AAPL_data = read.table("AAPL.csv", header=T, sep=',')
# AAPL_Adj_Close = zoo(AAPL_data$AAPL.Adjusted, as.Date(as.character(AAPL_data$Index), format = c("%Y-%m-%d")))
(autoplot(AAPL$AAPL.Adjusted,
          main  = "AAPL Adjusted Close from 2002-02-01 to 2017-01-31")
        + ylab("Adj Close")
        + xlab("Date"))
AAPL.Weekly = ts(Ad(to.weekly(AAPL)), frequency = 52)
autoplot(stl(AAPL.Weekly[,1], s.window="period"),
         main  = "STL Decomposition of Weekly Data",
         xlab  = "Date")
ggtsdisplay(AAPL$AAPL.Adjusted,
            main  = "AAPL Adjusted Close from 2002-02-01 to 2017-01-31",
            ylab  = "Adj Close",
            xlab  = "Date")
adf.test(AAPL$AAPL.Adjusted)
```
```{r}
AAPL.Adjusted_Log_Return = diff(log(AAPL$AAPL.Adjusted))*100
AAPL.Adjusted_Log_Return = AAPL.Adjusted_Log_Return[!is.na(AAPL.Adjusted_Log_Return)] # Remove missing values
(autoplot(AAPL.Adjusted_Log_Return,
          main  = sprintf("Log-Returns (×100%%)", 0))
        + ylab("Adj Close")
        + xlab("Date")
        + geom_abline(slope = 0, intercept = 0, col = 'red'))
AAPL.basicStats = basicStats(AAPL.Adjusted_Log_Return)
AAPL.basicStats
```
The average log return equals `r AAPL.basicStats[c("Mean"),]` with a standard error of `r AAPL.basicStats[c("SE Mean"),]`. Thus the mean of the return process if not statistically significantly different from zero.
```{r}
ggtsdisplay(AAPL.Adjusted_Log_Return,
            main  = "Log-Returns (×100%%)",
            ylab  = "Adj Close",
            xlab  = "Date")
ggtsdisplay(abs(AAPL.Adjusted_Log_Return),
            main  = "Absolute Log-Returns",
            ylab  = "Adj Close",
            xlab  = "Date")
ggtsdisplay(AAPL.Adjusted_Log_Return^2,
            main  = "Squared Log-Returns",
            ylab  = "Adj Close",
            xlab  = "Date")
```
```{r}
qqnorm(AAPL.Adjusted_Log_Return)
qqline(AAPL.Adjusted_Log_Return, col = 'red')
# kurtosis(AAPL.Adjusted_Log_Return)
# skewness(AAPL.Adjusted_Log_Return)
AAPL.basicStats[c("Kurtosis"),]
AAPL.basicStats[c("Skewness"),]
```
The sample kurtosis of the log returns equals `r AAPL.basicStats[c("Kurtosis"),]`. The positive kurtosis indicates a heavy-tailed distribution.

The skewness of the log returns equals `r AAPL.basicStats[c("Skewness"),]`. The negative skewness indicates the the data is skewed to the left left i.e. left tail is longer than right tail.
```{r}
eacf(AAPL.Adjusted_Log_Return^2)
```
EACF for squared log-returns suggests a GARCH(1,1) model and a possible GARCH(2,2) model.
```{r}
eacf(abs(AAPL.Adjusted_Log_Return))
```
EACF for absolute log-returns convincingly suggests a GARCH(1,1) model.

```{r}
garch11 = garch(AAPL.Adjusted_Log_Return, order = c(1,1))
summary(garch11)
AIC(garch11)
```
```{r}
plot(residuals(garch11),type='h',ylab='Standardized Residuals')
qqnorm(residuals(garch11)); qqline(residuals(garch11), col = 2)
acf(residuals(garch11)^2, na.action = na.omit)
gBox(garch11, method='squared')
```

## sGARCH(1,1)
### sGARCH(1,1), Normal Distribution (Innovations)
```{r}
ugarchfit(spec = ugarchspec(
  variance.model=list(garchOrder=c(1,1)),
  mean.model=list(armaOrder = c(0,0))),
  data = AAPL.Adjusted_Log_Return)
```
### sGARCH(1,1), Skew Normal Distribution
```{r}
ugarchfit(spec = ugarchspec(
  variance.model=list(garchOrder=c(1,1)),
  mean.model=list(armaOrder = c(0,0)),
  distribution.model = "snorm"),
  data = AAPL.Adjusted_Log_Return)
```
### sGARCH(1,1), T-Distribution
```{r}
ugarchfit(spec = ugarchspec(
  variance.model=list(garchOrder=c(1,1)),
  mean.model=list(armaOrder = c(0,0)),
  distribution.model = "std"),
  data = AAPL.Adjusted_Log_Return)
```
### sGARCH(1,1), Skew T-Distribution
```{r}
ugarchfit(spec = ugarchspec(
  variance.model=list(garchOrder=c(1,1)),
  mean.model=list(armaOrder = c(0,0)),
  distribution.model = "sstd"),
  data = AAPL.Adjusted_Log_Return)
```
### sGARCH(1,1), Generalized Error Distribution
```{r}
ugarchfit(spec = ugarchspec(
  variance.model=list(garchOrder=c(1,1)),
  mean.model=list(armaOrder = c(0,0)),
  distribution.model = "ged"),
  data = AAPL.Adjusted_Log_Return)
```
### sGARCH(1,1), Skew Generalized Error Distribution
```{r}
ugarchfit(spec = ugarchspec(
  variance.model=list(garchOrder=c(1,1)),
  mean.model=list(armaOrder = c(0,0)),
  distribution.model = "sged"),
  data = AAPL.Adjusted_Log_Return)
```
### sGARCH(1,1), Normal Inverse Gaussian Distribution
```{r}
ugarchfit(spec = ugarchspec(
  variance.model=list(garchOrder=c(1,1)),
  mean.model=list(armaOrder = c(0,0)),
  distribution.model = "nig"),
  data = AAPL.Adjusted_Log_Return)
```
### sGARCH(1,1), Generalized Hyperbolic Distribution
```{r}
ugarchfit(spec = ugarchspec(
  variance.model=list(garchOrder=c(1,1)),
  mean.model=list(armaOrder = c(0,0)),
  distribution.model = "ghyp"),
  data = AAPL.Adjusted_Log_Return)
```
### sGARCH(1,1), Johnson's S<sub>U</sub> Distribution
```{r}
ugarchfit(spec = ugarchspec(
  variance.model=list(garchOrder=c(1,1)),
  mean.model=list(armaOrder = c(0,0)),
  distribution.model = "jsu"),
  data = AAPL.Adjusted_Log_Return)
```
## T-Distribution
### fGARCH(1,1), GARCH, T-Distribution
```{r}
ugarchfit(spec = ugarchspec(
  variance.model=list(model = "fGARCH",
                      submodel = "GARCH",
                      garchOrder=c(1,1)),
  mean.model=list(armaOrder = c(0,0)),
  distribution.model = "std"),
  data = AAPL.Adjusted_Log_Return)
```
### fGARCH(1,1), TGARCH, T-Distribution
```{r}
ugarchfit(spec = ugarchspec(
  variance.model=list(model = "fGARCH",
                      submodel = "TGARCH",
                      garchOrder=c(1,1)),
  mean.model=list(armaOrder = c(0,0)),
  distribution.model = "std"),
  data = AAPL.Adjusted_Log_Return)
```
### fGARCH(1,1), AVGARCH, T-Distribution
```{r}
ugarchfit(spec = ugarchspec(
  variance.model=list(model = "fGARCH",
                      submodel = "AVGARCH",
                      garchOrder=c(1,1)),
  mean.model=list(armaOrder = c(0,0)),
  distribution.model = "std"),
  data = AAPL.Adjusted_Log_Return)
```
### fGARCH(1,1), NGARCH, T-Distribution
```{r}
ugarchfit(spec = ugarchspec(
  variance.model=list(model = "fGARCH",
                      submodel = "NGARCH",
                      garchOrder=c(1,1)),
  mean.model=list(armaOrder = c(0,0)),
  distribution.model = "std"),
  data = AAPL.Adjusted_Log_Return)
```
### fGARCH(1,1), NAGARCH, T-Distribution
```{r}
ugarchfit(spec = ugarchspec(
  variance.model=list(model = "fGARCH",
                      submodel = "NAGARCH",
                      garchOrder=c(1,1)),
  mean.model=list(armaOrder = c(0,0)),
  distribution.model = "std"),
  data = AAPL.Adjusted_Log_Return)
```
### fGARCH(1,1), APARCH, T-Distribution
```{r}
ugarchfit(spec = ugarchspec(
  variance.model=list(model = "fGARCH",
                      submodel = "APARCH",
                      garchOrder=c(1,1)),
  mean.model=list(armaOrder = c(0,0)),
  distribution.model = "std"),
  data = AAPL.Adjusted_Log_Return)
```
### fGARCH(1,1), GJRGARCH, T-Distribution
```{r}
ugarchfit(spec = ugarchspec(
  variance.model=list(model = "fGARCH",
                      submodel = "GJRGARCH",
                      garchOrder=c(1,1)),
  mean.model=list(armaOrder = c(0,0)),
  distribution.model = "std"),
  data = AAPL.Adjusted_Log_Return)
```
### fGARCH(1,1), ALLGARCH, T-Distribution
```{r}
ugarchfit(spec = ugarchspec(
  variance.model=list(model = "fGARCH",
                      submodel = "ALLGARCH",
                      garchOrder=c(1,1)),
  mean.model=list(armaOrder = c(0,0)),
  distribution.model = "std"),
  data = AAPL.Adjusted_Log_Return)
```
### eGARCH(1,1), T-Distribution
```{r}
ugarchfit(spec = ugarchspec(
  variance.model=list(model = "eGARCH",
                      garchOrder=c(1,1)),
  mean.model=list(armaOrder = c(0,0)),
  distribution.model = "std"),
  data = AAPL.Adjusted_Log_Return)
```
### gjrGARCH(1,1), T-Distribution
```{r}
ugarchfit(spec = ugarchspec(
  variance.model=list(model = "gjrGARCH",
                      garchOrder=c(1,1)),
  mean.model=list(armaOrder = c(0,0)),
  distribution.model = "std"),
  data = AAPL.Adjusted_Log_Return)
```
### apARCH(1,1), T-Distribution
```{r}
ugarchfit(spec = ugarchspec(
  variance.model=list(model = "apARCH",
                      garchOrder=c(1,1)),
  mean.model=list(armaOrder = c(0,0)),
  distribution.model = "std"),
  data = AAPL.Adjusted_Log_Return)
```
### iGARCH(1,1), T-Distribution
```{r}
ugarchfit(spec = ugarchspec(
  variance.model=list(model = "iGARCH",
                      garchOrder=c(1,1)),
  mean.model=list(armaOrder = c(0,0)),
  distribution.model = "std"),
  data = AAPL.Adjusted_Log_Return)
```
### csGARCH(1,1), T-Distribution
```{r}
ugarchfit(spec = ugarchspec(
  variance.model=list(model = "csGARCH",
                      garchOrder=c(1,1)),
  mean.model=list(armaOrder = c(0,0)),
  distribution.model = "std"),
  data = AAPL.Adjusted_Log_Return)
```
## Forecasting with eGARCH(1,1), T-Distribution
```{r}
egarch11 = ugarchfit(spec = ugarchspec(
  variance.model=list(model = "eGARCH",
                      garchOrder=c(1,1)),
  mean.model=list(armaOrder = c(0,0)),
  distribution.model = "std"),
  data = AAPL.Adjusted_Log_Return)

egarch11.forecast = ugarchforecast(egarch11, n.ahead = 30)
egarch11.forecast
plot(egarch11.forecast, which = 1)
plot(egarch11.forecast, which = 3)
```

```{r}
test_size = 90

egarch11 = ugarchfit(spec = ugarchspec(
  variance.model=list(model = "eGARCH",
                      garchOrder=c(1,1)),
  mean.model=list(armaOrder = c(0,0)),
  distribution.model = "std"),
  data = AAPL.Adjusted_Log_Return,
  out.sample = test_size)

egarch11.forecast = ugarchforecast(egarch11, n.ahead = test_size + 30, n.roll = test_size)
egarch11.forecast
plot(egarch11.forecast, which = 1)
plot(egarch11.forecast, which = 2)
plot(egarch11.forecast, which = 3)
plot(egarch11.forecast, which = 4)
```
```{r}
egarch11.forecast = ugarchroll(spec = ugarchspec(
  variance.model=list(model = "eGARCH",
                      garchOrder=c(1,1)),
  mean.model=list(armaOrder = c(0,0)),
  distribution.model = "std"),
  data = AAPL.Adjusted_Log_Return,
  n.ahead = 1,
  forecast.length = 90,
  refit.window = "recursive",
  refit.every = 30)

egarch11.forecast
plot(egarch11.forecast, which = 1)
plot(egarch11.forecast, which = 2)
plot(egarch11.forecast, which = 3)
plot(egarch11.forecast, which = 4)
plot(egarch11.forecast, which = 5)
```